{\fontfamily{ptm}\selectfont The Proximal Point interpretation of Nesterov accelerated proximal gradient}

### 2024-08-14 Notebook LaTEX/Momentum Accelerations/paper.tex
/w\, fixes and changes

    \begin{lemma}[Linearized Gradient Mapping Lower Bound]
    \label{lemma:grad_map_linearization}
        Continue from 
        \hyperref[lemma:grad_map_lemma_first]{definition \ref*{lemma:grad_map_lemma_first}}, 
        with $x^+ = \mathcal T_L(x)$, the gradient mapping satisfies the inequality for all $z$: 
        \begin{align*}
            h(z) &\ge
            h(z) - \frac{L}{2}\Vert x - x^+\Vert^2
            \ge h(x^+) + \langle \mathcal G_L(x), z - x\rangle. 
        \end{align*}
    \end{lemma}
    \begin{proof}
        Directly from the $L$-smoothness of $f$, convexity of $g, f$, we have the list of inequalities: 
        \begin{align*}
            &f(x^+) \le 
            f(x) + \langle \nabla f(x), x^+ - x\rangle
            + \frac{L}{2}\Vert x - x^+\Vert^2, 
            \\
            &f(x) + \langle \nabla f(x), z - x\rangle 
            \le f(z), 
            \\
            &g(x^+) \le 
            g(z) + \langle \partial g(x^+), x^+ - z\rangle. 
        \end{align*}
        Now, consider adding $g(x^+)$ to the first inequality from above we get 
        {\footnotesize 
        \begin{align*}
            f(x^+) + g(x^+) 
            &\le 
            f(x) + g(x^+) + \langle \nabla f(x), x^+ - x\rangle 
            + \frac{L}{2}\Vert x - x^+\Vert^2
            \\
            &\le 
            (f(z) - \langle \nabla f(x), z - x\rangle) + 
            \left(g(z) - \langle \partial g(x^+), x^+ - z\rangle\right)
            + 
            \langle \nabla f(x), x^+ - x\rangle + \frac{L}{2}\Vert x - x^+\Vert^2
            \\
            &= f(z) + g(z) + \langle \nabla f(x), x - z + x^+ - x\rangle
            + 
            \langle \partial g(x^+), x^+ - z\rangle + \frac{L}{2}\Vert x - x^+\Vert^2
            \\
            &= 
            h(z) + \langle \nabla f(x), x^+ - z\rangle + 
            \langle \partial g(x^+), x^+ - z\rangle
            + \frac{L}{2}\Vert x - x^+\Vert^2
            \\
            &= h(z) + \langle \nabla f(x) + \partial g(x^+), x^+ - z\rangle 
            + \frac{L}{2}\Vert x - x^+\Vert^2. 
        \end{align*}
        }
        By $\mathcal G_L(x) = L(x - x^+) \in \nabla f(x) + \partial g(x^+)$ from previous dicussion, we have 
        \begin{align*}
            h(x^+) &\le 
            h(z) + \langle \mathcal G_L(x), x^+ - z\rangle + \frac{L}{2}\Vert x - x^+\Vert^2
            \\
            &= h(z) - \langle L(x^+ - x), x^+ - x + x - z \rangle 
            + 
            \frac{L}{2}\Vert x - x^+\Vert^2
            \\
            &= 
            h(z) - L\Vert x^+ - x\Vert^2 
            + L \langle x^+ - x, x - z\rangle
            + \frac{L}{2}\Vert x - x^+\Vert^2
            \\
            &= h(z) + \langle \mathcal G_L(x), x - z\rangle - 
            \frac{L}{2}\Vert x - x^+\Vert^2. 
        \end{align*}
        Therefore, the inequality is justified. 
    \end{proof}
    \begin{remark}
        Observe that the linearization $h(x^+) + \langle \mathcal G_L(x), z - x\rangle$ is anchored at $x^+$, instead of $x$. 
        Geometrically, it's tilted and it ``prefers" the sharp corners of a convex function, if, $x$ is close to a sharp corner. 
    \end{remark}

\section{The PPM interpreation of accelerated gradient}\label{sec:ppm_interp_of_ag}
    In this section, we state the PPM formulation of the accelerated proximal gradient method. 
    The formulation is abstract because the parameters are not yet determined. 
    \begin{definition}[AG Proximal Gradient PPM Generic Form]
    \label{def:ag_prox_grad_ppm}
        Let $h=f + g$ be the sum of convex function $g$ and convex differentiable $f$ with $L$-Lipschitz gradient. 
        With $\mathcal T_L, \mathcal G_L$ being the proximal gradient, and the gradient mapping operator. 
        Define the linear lower bounding function for $h$ at $y$, for all $x$: 
        $$
        \begin{aligned}
            l_h(x; y) &= h(\mathcal T_L y) + \langle \mathcal G_L(y), x - y \rangle \le h(x), 
        \end{aligned}
        $$
        with that we define the algorithm:
        $$
        \begin{aligned}
            x_{t + 1} &= \argmin_{x} \left\lbrace
                l_h(x; y_t) + \frac{1}{2\tilde \eta_{t + 1}} 
                \Vert x - x_t\Vert^2
            \right\rbrace,
            \\
            y_{t + 1}&= 
            \argmin_{x}
            \left\lbrace
                l_h(x; y_t) + \frac{L}{2}\Vert x - y_t\Vert^2 + 
                \frac{1}{2\eta_{t + 1}} \Vert x - x_{t + 1}\Vert^2
            \right\rbrace.
        \end{aligned}
        $$
    \end{definition}

@@

>>>>

    \begin{lemma}[Linearized Gradient Mapping Lower Bound]
    \label{lemma:grad_map_linearization}
        Continue from 
        \hyperref[lemma:grad_map_lemma_first]{definition \ref*{lemma:grad_map_lemma_first}}, 
        further assume that $f$ is strongly convex with index $\mu \ge 0$,
        with $x^+ = \mathcal T_L(x)$, the gradient mapping satisfies the inequality for all $z$: 
            h(z) &\ge 
            h(x^+) + \langle \mathcal G_L (x), z - x\rangle + 
            \frac{L}{2}\Vert x - x^+\Vert^2 + \frac{\mu}{2}
            \Vert z - x\Vert^2. 
        Directly from the $L$-smoothness of $f$, convexity of $g, f$, we have the list of inequalities: 
            &f(x^+) \le 
            f(x) + \langle \nabla f(x), x^+ - x\rangle
            + \frac{L}{2}\Vert x - x^+\Vert^2, 
            \frac{\mu}{2}\Vert z - x\Vert^2+ 
            f(x) + \langle \nabla f(x), z - x\rangle 
            \le f(z), 
            &g(x^+) \le 
            g(z) + \langle \partial g(x^+), x^+ - z\rangle. 
        Then we have the following sequence of relations 
            h(x^+) &= f(x^+) + g(x^+)
            \\&
            \begin{aligned}
                    f(x) + \langle \nabla f(x), x^+ - x\rangle
                    + \frac{L}{2}\Vert x - x^+\Vert^2
                &\qquad  
                + (g(z) + \langle \partial g(x^+), x^+ - z\rangle)
            \end{aligned}
            \\&
            \begin{aligned}
                    f(z) - \langle \nabla f(x), z - x\rangle
                    - \frac{\mu}{2}\Vert z - x\Vert^2
                    + \langle \nabla f(x), x^+ - x\rangle
                    \frac{L}{2}\Vert x - x^+\Vert^2
                &\qquad 
                (g(z) + \langle \partial g(x^+), x^+ - z\rangle)
            \end{aligned}
            \\&
            \begin{aligned}
                (f(z) + h(z)) 
                &\qquad 
                + \left(
                    \langle \nabla f(x), x - z\rangle + 
                    \langle \nabla f(x), x^+ - x\rangle + 
                    \langle \partial g(x^+), x^+ - z\rangle
                &\qquad 
                - \frac{\mu}{2}\Vert z - x\Vert^2
                + \frac{L}{2}\Vert x - x^+\Vert^2
            \end{aligned}
            \\& 
            \begin{aligned}
                &= h(z) + 
                    \langle \nabla f(x), x - x^+ + x^+ - z\rangle + 
                    \langle \nabla f(x), x^+ - x\rangle + 
                    \langle \partial g(x^+), x^+ - z\rangle
                &\qquad 
                - \frac{\mu}{2}\Vert z - x\Vert^2
                + \frac{L}{2}\Vert x - x^+\Vert^2
            \end{aligned}
            \\& 
            \begin{aligned}
                &= h(z) + 
                \langle \nabla f(x) + \partial g(x^+), x^+ - z\rangle 
                - \frac{\mu}{2}\Vert z - x\Vert^2
                + \frac{L}{2}\Vert x - x^+\Vert^2
            \end{aligned}
            \\& 
            \begin{aligned}
                &= h(z) + 
                    \langle \mathcal G_L(x), x^+ - z\rangle 
                - \frac{\mu}{2}\Vert z - x\Vert^2
                + \frac{L}{2}\Vert x - x^+\Vert^2
            \end{aligned}
            \\& 
            \begin{aligned}
                &= h(z) + \langle L(x - x^+), x^+ - x + x - z\rangle 
                - \frac{\mu}{2}\Vert z - x\Vert^2
                + \frac{L}{2}\Vert x - x^+\Vert^2
            \end{aligned}
            \\&
            \begin{aligned}
                &= h(z) + 
                \underbrace{\langle L(x - x^+), x - z\rangle}_{
                    = - \langle \mathcal G_L (x), z - x\rangle
                - \frac{\mu}{2}\Vert z - x\Vert^2
                - \frac{L}{2}\Vert x - x^+\Vert^2
            \end{aligned}. 
        Moving everything except $h(z)$ from the RHS to the LHS yield the desired inequality. 
        Observe that the linearization $h(x^+) + \langle \mathcal G_L(x), z - x\rangle$ is anchored at $x^+$, instead of $x$. 
        Geometrically, it's tilted and it ``prefers" the sharp corners of a convex function, if, $x$ is close to a sharp corner.

\section{The PPM interpreation of accelerated gradient}\label{sec:ppm_interp_of_ag}
    In this section, we state the PPM formulation of the accelerated proximal gradient method. 
    The formulation is abstract because the parameters are not yet determined. 
    \begin{definition}[AG Proximal Gradient PPM Generic Form]
    \label{def:ag_prox_grad_ppm}
        Let $h=f + g$ be the sum of convex function $g$ and convex differentiable $f$ with $L$-Lipschitz gradient. 
        With $\mathcal T_L, \mathcal G_L$ being the proximal gradient, and the gradient mapping operator. 
        Define the linear lower bounding function for $h$ at $y$, for all $x$: 
        $$
        \begin{aligned}
            l_h(x; y) &= h(\mathcal T_L y) + \langle \mathcal G_L(y), x - y \rangle 
                + \frac{L}{2}\Vert y - \mathcal T_L y\Vert^2
            \le h(x), 
        \end{aligned}
        $$
        with that we define the algorithm:
        $$
        \begin{aligned}
            x_{t + 1} &= \argmin_{x} \left\lbrace
                l_h(x; y_t) + \frac{1}{2\tilde \eta_{t + 1}} 
                \Vert x - x_t\Vert^2
            \right\rbrace,
            \\
            y_{t + 1}&= 
            \argmin_{x}
            \left\lbrace
                l_h(x; y_t) + \frac{L}{2}\Vert x - y_t\Vert^2 + 
                \frac{1}{2\eta_{t + 1}} \Vert x - x_{t + 1}\Vert^2
            \right\rbrace.
        \end{aligned}
        $$
    \end{definition}

@@

        It will be used next. 
